---
title: "Sampling distribution Bootstrapping"
author: "Jacob Cathcart"
date: "6/27/2018"
output: html_document
---

```{r}
library(dplyr)
library(car)

X <- runif(100000, 0, 1)

Y <- runif(100000, 0, 1)

Q <- runif(100000, 0, 1)

A <- X + Y

B <- X +Y + Q

```


# What is the distribution of A/B and the probability A/B > 1.2?

```{r}

hist(A)
hist(B)


mean(A)
var(A)
mean( A > 1.2)


mean(B)
var(B)
mean(B > 1.2)

```


# What are the boundarires of the top 5% and the middle 95%

```{r}
quantile(B, .95)

quantile(B,c(.025,.975))

```


# What is the distribution of X-bar (X1, X2, X3 ~ iid UNIF(0,1))

```{r}
N <- 10^5
n <- 3

simmeans <- numeric(N)

for(i in 1:N){
  simsample <- runif(n, 0, 1)
  simmeans[i] <- mean(simsample)
}

hist(simmeans)
mean(simmeans)
var(simmeans)

qqPlot(simmeans)


```
##The shape is not normal here so we cannot use CLT or Pnorm 

##So,


```{r}
mean(simmeans >= 0.7)
```

## What if I tried to use pnorm incorrectly?

```{r}
pnorm(.7,.5, sqrt(1/36), F)
```

## What is the distribution of X-bar (X1, X2, X3 ~ iid EXP(0,1))

```{r}
N <- 10^5
n <- 3 

simmeans <- numeric(N)

for(i in 1:N){
  simsample <- rexp(n, 1)
  simmeans[i] <- mean(simsample)
}

hist(simmeans)
mean(simmeans)
var(simmeans)


qqPlot(simmeans)

```

# IDK what to call this

```{r}
mlb <- read.csv("https://raw.githubusercontent.com/STAT-JET-ASU/Datasets/master/Instructor/mlb2002.csv") %>% 
  mutate( salary = salary /1000000)
mean(mlb$salary)
sd(mlb$salary)

sampledat <- sample(mlb$salary, 10)
mean <- mean(sampledat)
sd <- sd(sampledat)

print(mean)

mean + 1.96*sd/sqrt(10)
mean  - 1.96*sd/sqrt(10)


qt(.025,9)


```

# Bootstraping fish examples

```{r}
FISH <- read.csv("https://raw.githubusercontent.com/STAT-JET-ASU/Datasets/master/Chihara/Walleye.csv")

str(FISH)
summary(FISH)

boxplot(FISH$Length)

inchesmean <- mean(FISH$Length)
inchesstdv <- sd(FISH$Length)
inches_n   <- length(FISH$Length)
inches_df  <- inches_n - 1

# what if we do a traditional 95% CI?

  CI_LB <- inchesmean - qt(0.975, inches_df) * inchesstdv / sqrt(inches_n)
  CI_UB <- inchesmean + qt(0.975, inches_df) * inchesstdv / sqrt(inches_n)
  print(c(CI_LB, CI_UB))

# bootstrap 95% percentile interval
  
N <- 10^4
bootmeans <- numeric(N)
for (i in 1:N){
  bootsample   <- sample(FISH$Length, inches_n, replace = TRUE)
  bootmeans[i] <- mean(bootsample)
}

quantile(bootmeans, c(0.025, 0.975))

# is the distribution approximately normal?
hist(bootmeans)
require(car)
qqPlot(bootmeans)
```

## Same thing but with weight

```{r}


boxplot(FISH$Weight)


weightmean <- mean(FISH$Weight)
weightstdv <- sd(FISH$Weight)
weight_n   <- length(FISH$Weight)
weight_df  <- weight_n - 1

# traditional 95% CI

CI_LB <- weightmean - qt(0.975, weight_df) * weightstdv / sqrt(weight_n)
CI_UB <- weightmean + qt(0.975, weight_df) * weightstdv / sqrt(weight_n)
print(c(CI_LB, CI_UB))

# bootstrap 95% percentile interval

N <- 10^4
bootmeans <- numeric(N)
for (i in 1:N){
  bootsample   <- sample(FISH$Weight, weight_n, replace = TRUE)
  bootmeans[i] <- mean(bootsample)
}

quantile(bootmeans, c(0.025, 0.975))

# is the distribution approximately normal?
hist(bootmeans)
require(car)
qqPlot(bootmeans)


```

