<hr>

Read in each dataset using the `read.csv` function and use the following functions to explore its structure. You can find all the data on our [class dataset list](https://stat-jet-asu.github.io/Datasets/DatasetList.html). The flight delays and Verizon datasets are from the textbook, the rest are instructor datasets.

* `str`
* `glimpse`
* `summary`
* `head` to show three rows of the datset
* `sum` and `complete.cases` to find the number of rows that contain NA's

```{r, message=FALSE, warning=FALSE}
require(dplyr)
```


### Annual Inflation Data

```{r}
inflation <- read.csv("https://raw.githubusercontent.com/STAT-JET-ASU/Datasets/master/Instructor/annualinflation.csv")
str(inflation)
glimpse(inflation)
summary(inflation)
head(inflation)
sum(complete.cases(inflation))
```


### Anthropometric Data

```{r}
anthro <- read.csv("https://raw.githubusercontent.com/STAT-JET-ASU/Datasets/master/Instructor/anthropometric.csv")
str(anthro)
glimpse(anthro)
summary(anthro)
head(anthro)
sum(complete.cases(anthro))
```


### Flight Delays Data

```{r}
flightdelays <- read.csv("https://raw.githubusercontent.com/STAT-JET-ASU/Datasets/master/Chihara/FlightDelays.csv")
str(flightdelays)
glimpse(flightdelays)
summary(flightdelays)
head(flightdelays)
sum(complete.cases(flightdelays))
```


### Mauna Loa CO2 Data

```{r}
maunaloa <- read.csv("https://raw.githubusercontent.com/STAT-JET-ASU/Datasets/master/Instructor/maunaloaCO2.csv")
str(maunaloa)
glimpse(maunaloa)
summary(maunaloa)
head(maunaloa)
sum(complete.cases(maunaloa))
```


### STT3850 Personality Data

```{r}
personality <- read.csv("https://raw.githubusercontent.com/STAT-JET-ASU/Datasets/master/Instructor/personalitySTT3850.csv")
str(personality)
glimpse(personality)
summary(personality)
head(personality)
sum(complete.cases(personality))
```


### Popular Kids Data

```{r} 
popularkids <- read.csv("https://raw.githubusercontent.com/STAT-JET-ASU/Datasets/master/Instructor/popular.csv")
str(popularkids)
glimpse(popularkids)
summary(popularkids)
head(popularkids)
sum(complete.cases(popularkids))
```


### Powerball Jackpot Data

```{r}
loto <- read.csv("https://raw.githubusercontent.com/STAT-JET-ASU/Datasets/master/Instructor/powerballjackpot.csv")
str(loto)
glimpse(loto)
summary(loto)
head(loto)
sum(complete.cases(loto))
```


### Smallpox Data

```{r}
smallpox <- read.csv("https://raw.githubusercontent.com/STAT-JET-ASU/Datasets/master/Instructor/smallpox.csv")
str(smallpox)
glimpse(smallpox)
summary(smallpox)
head(smallpox)
sum(complete.cases(smallpox))
```


### Verizon Repair Times Data

```{r}
verizonrepairtime <- read.csv("https://raw.githubusercontent.com/STAT-JET-ASU/Datasets/master/Chihara/Verizon.csv")
str(verizonrepairtime)
glimpse(verizonrepairtime)
summary(verizonrepairtime)
head(verizonrepairtime)
sum(complete.cases(verizonrepairtime))
```


### Watauga County Schools Weather Data

```{r}
weather <- read.csv("https://raw.githubusercontent.com/STAT-JET-ASU/Datasets/master/Instructor/wcsweatherdays.csv")
str(weather)
glimpse(weather)
summary(weather)
head(weather)
sum(complete.cases(weather))
```



