<hr>

```{r, message=FALSE, warning=FALSE}
library(dplyr)
library(ggplot2)

draftdata <- read.csv("https://raw.githubusercontent.com/STAT-JET-ASU/Datasets/master/Instructor/vietnamdraft.csv")
```

### The Problem

To understand whether or not the 1970 Vietnam draft was random, we need to know what a random assignment would look like. Technically, any ordering (permutaion) of the 366 draft numbers could be produced by a random draw. Certain combinations, however, are less likely than others when we take into account grouping variables. We will use the `sample` function often for this purpose. 


### Performing Our Own Random Draft Assignment

```{r}
draftnumbers <- 1:366
draftdata$ourdraft <- sample(draftnumbers)

str(draftdata)
glimpse(draftdata)
summary(draftdata)
```


```{r}
ggplot(draftdata, aes(x = ourdraft)) + 
  geom_histogram(aes(y=..density..), breaks = seq(from = 0, to = 366), fill = "lightblue") +
  ggtitle("Discrete Uniform Distribution 1 to 366, probability = 1/366")  +
  xlab("draft number") +
  ylab("probability")

ggplot(draftdata, aes(x = ourdraft)) + 
  geom_density(fill = "lightblue") +
  ggtitle("Smoothed Density Plot of Draft Numbers")

summaryH <- draftdata %>% 
  group_by(halfyear) %>%
  summarize(observations = n(),
            average      = mean(ourdraft),
            standarddev  = sd(ourdraft),
            minimum      = min(ourdraft),
            maximum      = max(ourdraft))

print.data.frame(summaryH)

ggplot(draftdata, aes(x = ourdraft, fill = halfyear)) + 
  geom_density(alpha = 0.3) +
  scale_fill_manual(values = c("red", "yellow")) +
  ggtitle("Smoothed Density Plots of Draft Numbers for Each Half-Year")
```


```{r}
summaryM <- draftdata %>% 
  group_by(month) %>%
  summarize(observations = n(),
            average      = mean(ourdraft),
            standarddev  = sd(ourdraft),
            minimum      = min(ourdraft),
            maximum      = max(ourdraft))

print.data.frame(summaryM)

ggplot(summaryM, aes(x = month, y = average)) +
  geom_bar(stat = "identity", fill = "lightblue") +
  geom_errorbar(aes(ymin = average - standarddev/sqrt(observations), 
                    ymax = average + standarddev/sqrt(observations)), 
                width=.1) +
  ggtitle("Mean Draft Number by Month with Standard Error") +
  ylab("mean +/- SE")

ggplot(draftdata, aes(x = month, y = ourdraft)) + 
  geom_boxplot(fill = "lightgreen") +
  ggtitle("Boxplots of Draft Numbers by Month for 1970") +
  ylab("draft number")

ggplot(draftdata, aes(x = 1:366, y = ourdraft)) +
  geom_point() +
  geom_smooth() +
  xlab("order by birthday") +
  ylab("draft number")
```

